---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# https://community.rstudio.com/t/showing-only-the-first-few-lines-of-the-results-of-a-code-chunk/6963/2
#
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    lines <- options$output.lines
    if (is.null(lines)) {
      return(hook_output(x, options))  # pass to default hook
    }
    x <- unlist(strsplit(x, "\n"))
    more <- "..."
    if (length(lines)==1) {        # first n lines
      if (length(x) > lines) {
        # truncate the output, but add ....
        x <- c(head(x, lines), more)
      }
    } else {
      x <- c(more, x[lines], more)
    }
    # paste these lines together
    x <- paste(c(x, ""), collapse = "\n")
    hook_output(x, options)
  }
)
```

# steward

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of **steward** is to make it easier to import, manage, and publish the metadata associated with data frames. This might be useful for:

- [documenting a dataset](https://r-pkgs.org/data.html#documenting-data) for a package
- publishing a data dictionary in an R Markdown document

Our current defintion of metadata includes the name, description and source of a dataset, as well as the name, type, and description of each of the variables in the dataset.

The name, steward, is an homage to the [Data Stewardship](http://agron590-isu.github.io/) class taught by [Andee Kaplan](https://github.com/andeek) and [Ranae Dietzel](https://github.com/ranae) (also an author of this package) at Iowa State University in Fall 2016.

## Installation

You can install the development version of steward from [GitHub](https://github.com/uncoast-unconf/steward) with:

```r
# install.packages("devtools")
devtools::install_github("uncoast-unconf/steward")
```

## Usage

```{r load}
library("steward")
```

The current capabilities all deal with metadata:

- [read from YAML](#read-yaml) 
- [read from CSV](#read-csv) 
- [write to YAML](#write-yaml)
- [write to roxygen](#write-roxygen) 
- [write to gt table](#write-gt-table) 

### Future capabilities

Soon, we hope to add additional capabilities:

- [write metadata to YAML](https://github.com/uncoast-unconf/steward/issues/42)
- [combine metadata with dataset](https://github.com/uncoast-unconf/steward/issues/36)
  - [read and write combined dataset](https://github.com/uncoast-unconf/steward/issues/41) from/to flat files
- take into account [timezone as column metadata](https://github.com/uncoast-unconf/steward/issues/39)
- [build metadata from package-documentation](https://github.com/uncoast-unconf/steward/issues/43) (`.Rd` file)

#### Read YAML

Here are the first lines of a YAML file containing metadata from [ggplot2](http://ggplot2.tidyverse.org)'s `diamonds` dataset, which is available as:

```r
system.file("metadata/diamonds.yml", package = "steward")
```

```{r yaml, echo = FALSE, comment = "", output.lines = 11}
system.file("metadata/diamonds.yml", package = "steward") %>%
  readr::read_lines() %>%
  glue::glue_collapse(sep = "\n")
```

We can read a YAML file using `stw_read_yaml()`:

```{r example_read_yaml}
stw_read_yaml(system.file("metadata/diamonds.yml", package = "steward"))
```

#### Read CSV

This is an emerging capability that we need to document.

#### Write YAML

To get a YAML string from metadata, you can use the `stw_to_yaml()` function. Here are the first few lines:

```{r stw_to_yaml, output, output.lines = 11}
cat(stw_to_yaml(diamonds_meta))
```

Other functions in this family:

- `stw_use_yaml()` copies the YAML to the clipboard.
- `stw_write_yaml()` writes the YAML string to a file.

#### Write roxygen

```{r example_roxygen}
cat(stw_to_roxygen(diamonds_meta))
```

#### Write gt Table

If you are creating an R Markdown document, you can use the `stw_to_table()` function to create a [gt table](https://gt.rstudio.com):

```{r}
stw_to_table(diamonds_meta)
```

## Related work

The [codebook](https://rubenarslan.github.io/codebook/) package can help you manage dataset metadata.









