---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# steward

<!-- badges: start -->
<!-- badges: end -->

The goal of steward is to enable a package developer to embed a data dictionary within their package using the Roxygen syntax. his is especially useful during package development where the developer would like to embed an example dataset with a data dictionary. This results in package users being able to pull up the data dictionary using the helper function and view the appropriate help document.

In the past, the developer would have to code by hand using Roxygen syntax the data dictionary. This can lead to increased development time and frustration. However, the Steward package aims to take a data dictionary in either a YAML or CSV format and automatically create a new R file for the data dictionary contents implemented using the Roxygen syntax.

## Class
The steward package is equipped with its own S3 class that enables it to seamlessly read in either a YAML or CSV file and convert to a Roxygen syntax.

## Installation

You can install the development version of steward from [GitHub](https://github.com/uncoast-unconf/steward) with:

## Example: Read YAML

This is a basic example which shows you how to solve a common problem:

```{r example}
library(steward)
stw_read_yaml(system.file("metadata/diamonds.yaml", package = "steward"))
```
