---
title: "steward"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{steward}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = ""
)

# https://community.rstudio.com/t/showing-only-the-first-few-lines-of-the-results-of-a-code-chunk/6963/2
#
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    lines <- options$output.lines
    if (is.null(lines)) {
      return(hook_output(x, options))  # pass to default hook
    }
    x <- unlist(strsplit(x, "\n"))
    more <- "..."
    if (length(lines)==1) {        # first n lines
      if (length(x) > lines) {
        # truncate the output, but add ....
        x <- c(head(x, lines), more)
      }
    } else {
      x <- c(more, x[lines], more)
    }
    # paste these lines together
    x <- paste(c(x, ""), collapse = "\n")
    hook_output(x, options)
  }
)
```

```{r setup}
library("steward")
```

The goal of this package is to make it a little easier to build and publish data-dictionaries, in particular:

- for [package documentation](https://r-pkgs.org/data.html#documenting-data). 
- to produce [nicely-formatted](https://gt.rstudio.com) data-dictionaries in RMarkdown documents.

In this article, we will look at a few (what we think are) common tasks:

- [create dataset](#create) 
- [write dataset to package-documentation](#write-to-package)
- [write dataset to gt table](#write-gt-table) 

## Create 

For this set of examples, we use the **ggplot2** `diamonds` dataset. We will create an `stw_dataset`, which is a data frame with an extra class attached to help manage metadata. An `stw_dataset` **is** a data frame, in the same way that a [tibble](https://tibble.tidyverse.org) is a data frame. 

In the code that follows we have three steps:

1. Create the `stw_dataset()` using a data frame.
2. Add the top-level metadata using `stw_mutate_meta()`.
3. Add the column-level metadata using `stw_mutate_dict()`.

The functions `stw_mutate_meta()` and `stw_mutate_dict()` work a little bit like `dplyr::mutate()`. However, instead of mutating the columns of a data frame, `stw_mutate_meta()` mutates the top-level metadata, and `stw_mutate_dict()` mutates the *descriptions* of columns in a data frame:

```{r}
diamonds_new <-
  stw_dataset(ggplot2::diamonds) %>%
  stw_mutate_meta(
    name = "diamonds",
    title = "Prices of 50,000 round cut diamonds",
    description = "A dataset containing the prices and other attributes of almost 54,000 diamonds.",
    sources = list(
      list(
        title = "DiamondSearchEngine",
        path = "http://www.diamondse.info/"
      )
    )
  ) %>%
  stw_mutate_dict(
    price = "price in US dollars ($326--$18,823)",
    carat = "weight of diamond (0.2--5.01)",
    cut = "quality of the cut (Fair, Good, Very Good, Premium, Ideal)",
    color = "diamond color, from D (best) to J (worst)",
    clarity = "a measurement of how clear the diamond is (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best))",
    x = "length in mm (0--10.74)",
    y = "width in mm (0--58.9)",
    z = "depth in mm (0--31.8)",
    depth = "total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43--79)",
    table = "width of top of diamond relative to widest point (43--95)"
  )
```

We also offer a function to make some basic checks on the completeness of the metadata:

```{r}
stw_check(diamonds_new, verbosity = "all")
```

## Write to package

If you are including a dataset in a package, it is good practice to [document it](https://r-pkgs.org/data.html#documenting-data); in fact, CRAN insists!

You can use an `stw_dataset` to keep your documentation associated with your dataset as you build it for your package.
By keeping all the "stuff" together, this helps assure that the dataset documentation is complete and current.

You can use the function `stw_use_data()` wraps `usethis::use_data()` and `steward::stw_write_roxygen()`. For example, by running:

```r
stw_use_data(diamonds_new)
```

You will:

- validate that the metadata is complete.
- write the `diamonds_new` dataset to your package.
- write the roxygen documentation to `R/data-diamonds_new.R`.

## Create gt table

If you are creating an RMarkdown document, you can use the `stw_to_table()` function to create a [gt](https://gt.rstudio.com) table:

```{r}
stw_to_table(diamonds_new)
```

