---
title: "steward"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{steward}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = ""
)

# https://community.rstudio.com/t/showing-only-the-first-few-lines-of-the-results-of-a-code-chunk/6963/2
#
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    lines <- options$output.lines
    if (is.null(lines)) {
      return(hook_output(x, options))  # pass to default hook
    }
    x <- unlist(strsplit(x, "\n"))
    more <- "..."
    if (length(lines)==1) {        # first n lines
      if (length(x) > lines) {
        # truncate the output, but add ....
        x <- c(head(x, lines), more)
      }
    } else {
      x <- c(more, x[lines], more)
    }
    # paste these lines together
    x <- paste(c(x, ""), collapse = "\n")
    hook_output(x, options)
  }
)
```

```{r setup}
library("steward")
```

The current capabilities all deal with metadata:

- [read from YAML](#read-yaml) 
- [read from CSV](#read-csv) 
- [write to YAML](#write-yaml)
- [write to roxygen](#write-roxygen) 
- [write to gt table](#write-gt-table) 

#### Read YAML

Here are the first lines of a YAML file containing metadata from [ggplot2](http://ggplot2.tidyverse.org)'s `diamonds` dataset, which is available as:

```r
system.file("metadata/diamonds.yml", package = "steward")
```

```{r yaml, echo = FALSE, comment = "", output.lines = 11}
system.file("metadata/diamonds.yml", package = "steward") %>%
  readr::read_lines() %>%
  glue::glue_collapse(sep = "\n")
```

We can read a YAML file using `stw_read_meta_yaml()`:

```{r example_read_yaml}
stw_read_meta_yaml(system.file("metadata/diamonds.yml", package = "steward"))
```

#### Read CSV

This is an emerging capability that we need to document.

#### Write YAML

To get a YAML string from metadata, you can use the `stw_to_yaml()` function. Here are the first few lines:

```{r stw_to_yaml, output, output.lines = 11}
cat(stw_to_yaml(diamonds_meta))
```

Other functions in this family:

- `stw_use_yaml()` copies the YAML string to the clipboard.
- `stw_write_yaml()` writes the YAML string to a file.

#### Write roxygen

To get a roxygen string from metadata, you can use the `stw_to_roxygen()` function:


```{r example_roxygen}
cat(stw_to_roxygen(diamonds_meta))
```

Other functions in this family:

- `stw_use_roxygen()` copies the roxygen string to the clipboard.
- `stw_write_roxygen()` writes the roxygen string to a file.

#### Write gt Table

If you are creating an R Markdown document, you can use the `stw_to_table()` function to create a [gt table](https://gt.rstudio.com):

```{r}
stw_to_table(diamonds_meta)
```

